---
title: "data exploration and analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
```

## Questions

1- how many companies do we have

2- how many companies do we have per month

3- what is the company with the maximum/minimum market value  

4- what is the maxium , minimum  and average market value per month

5- company with the highest monthly returns () monthly returns: last day of the month / first day in the month - 1)

6-  top 3 firms we are interested in


```{r}
d_tidy <- read_rds("data/d_tidy.rds")

d_tidy
```


## number of firms

```{r number of firms}

d_tidy %>%
  distinct(company) %>%
  summarise(n = n())
  

```

## number of firms per month

```{r}
d_tidy %>%
  mutate(month = months(dates)) %>%
  group_by(month) %>%
  distinct(company) %>%
  summarise(n = n()) 
```


The company with the maximum/minimum market value  

```{r}
d_tidy %>%
  summarise(max_val = max(val, na.rm = TRUE),
            min_val = min(val, na.rm = TRUE))
```

```{r}
d_tidy %>%
  filter(val %in% max(val, na.rm = TRUE) |
         val %in% min(val, na.rm = TRUE))
```


## monthly min, max and average market value

```{r}
a <- d_tidy %>%
  group_by(dates) %>%
  summarise(daily_val = sum(val, na.rm = TRUE) ) %>%
  mutate(month = months(dates)) %>%
  group_by(month) %>%
  summarise(max_val = max(daily_val,na.rm = TRUE),
            min_val = min(daily_val, na.rm = TRUE),
            average = mean(daily_val, na.rm = TRUE),
            std = sd(daily_val, na.rm = TRUE)) %>%
  arrange(desc(average))
write_csv(a,"~/Desktop/a.csv")
  
```



## company with the highest monthly returns

```{r}
d_returns <-  
  d_tidy %>%
  mutate(month = months(dates)) %>%
  arrange(company, dates) %>%
  group_by(company,month) %>%
  slice(c(1,n())) %>%
  mutate(returns = round(val/shift(val)-1,2)) %>%
  ungroup() %>%
  na.omit() %>%
  select(-dates, -val, -company_code) 

d_returns
```

```{r}
d_returns %>%
  filter(returns %in% max(returns) |
        returns %in% min(returns))
```


## selected firms

```{r}
big_firms <- c("Jarir Marketin", "Almara","Saudi Basic Industrie")

d_big_firms <- 
  d_tidy %>%
  filter(company %in% big_firms )

d_big_firms
```


# visualizing their daily returns

```{r}

d_big_firms %>%
  ggplot(aes(x = dates, y = val))+
  geom_line()
```


```{r color by company}

d_big_firms %>%
  ggplot(aes(x = dates, y = val, col = company))+
  geom_line()
```


```{r facet}

d_big_firms %>%
  ggplot(aes(x = dates, y = val, col = company))+
  geom_line() +
  facet_wrap(~company, scales = "free_y")
```

```{r}
d_big_firms %>%
  ggplot(aes(x = dates, y = val, col = company))+
  geom_line() +
  facet_wrap(~company, scales = "free_y", ncol = 1)
```


```{r}
p <- d_big_firms %>%
  ggplot(aes(x = dates, y = val, col = company))+
  geom_line() +
  facet_wrap(~company, scales = "free_y", ncol = 1)+
  scale_y_continuous(label = scales::comma) + 
  labs(
    title = "Market value of big firms in the Saudi stock market",
    x = "",
    y = "market value"
  ) + 
  theme(legend.position="none")

p
```

## interactive plots


```{r}

library(plotly)

ggplotly(p)
```



# modeling


```{r}
d_tidy %>%
  mutate(month =lubridate::month(dates)) %>%
  group_by(company) %>%
  nest(month, val) %>% 
  mutate(reg_info  = map(data, function(x){
    fit <- lm(val~month, data = x)
    data.frame(
      intercept = fit$coefficients[1],
      slope = fit$coefficients[2]
    )
  })
) %>%
  select(-data) %>%
  unnest(reg_info) %>%
  arrange(desc(slope)) %>% View()
```

